---
title: "Example workflow qPCRanalysis"
author: "Jordi Camps"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output: rmarkdown::github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path = "README_figs/README-")
```

## Load packages
```{r message=FALSE}
library(readxl)
library(tidyr)
library(dplyr)
library(ggplot2)
library(qPCRanalysis)
library(ggpubr)
library(lazyeval)
```

## Load data
We load the data with the **read_excel()** function of the readxl package. The data should consist of at least 4 columns:

* __Sample__ : the biological variability in your data set like a disease, different timepoints, perturbation ...
* __Replicate__ : invidividual names per sample
* __Gene__ : the gene names of your primers
* __CT__ : the raw CT values

```{r message=FALSE, warning=FALSE}
qpcr <- read_excel("vignette/qPCR_data.xlsx", col_names = TRUE, sheet = 3, skip = 35)
head(qpcr)
```

Now we need to clean the data frame. We will select only the columns that we need and change their names because R does not like spaces inside column names.
```{r}
qpcr <- select(qpcr, c("Sample Name", "Target Name", "CT"))
colnames(qpcr) <- c("Sample", "Gene", "CT")
head(qpcr)
```

In this example we want to see the difference between different genotypes. We will split the Sample Name column in two.
```{r}
qpcr <- separate(qpcr, col = Sample, into = c("Sample", "Replicate"), sep = "_")
qpcr <- filter(qpcr, Sample != "Blanc")
head(qpcr)
```

When we check the structure of the data frame we see that the CT column is not numeric, we will change this. In addition, all undetermined values are switched to NA. If desired, these values can be changed to 40.
```{r}
qpcr$CT <- as.numeric(qpcr$CT)
str(qpcr)
```

## Calculate Delta CT
To calculate delta CT we use the `calculate_DCT()` function. This function requires four arguments:

* __df__ : dataframe structured like the proposed data file
* __hkg__ : name of housekeeping gene or genes that you want to use to normalize against
* __sample_col__ : name of the sample column
* __gene_col__ : name of the gene column

It will pass a dataframe with three added columns: 

* __CT_hkg__ : average CT value of housekeeping genes
* __DCT__ : Delta CT values
* __RE__ : relative expression to hkg

```{r}
qpcr <- calculate_DCT(df = qpcr, hkg = c("Rab35", "Rpl13a", "PSma3"), sample_col = "Sample", gene_col = "Gene")
head(qpcr)
```

## Statistics
Perform statistical test between two groups and add information to qpcr data frame. We will do this with the ggpubr package.
```{r}
stat <- compare_means(DCT ~ Sample, qpcr, method = "t.test", paired = FALSE, group.by = "Gene")
head(stat)
```

### Plot
```{r}
ggplot(qpcr, aes(x = Sample, y = DCT, col = Sample)) +
  geom_boxplot() +
  facet_wrap(~Gene, scales = "free_y") +
  stat_compare_means(method = "t.test", label = "p.signif", label.x = 1.5)
```

## Calculate Delta Delta CT
To calculate Delta Delta CT use the `calculate_DDCT()` function. This function  can only be run after the `calculate_DCT()` function is used and requires five argeuments:

* __df__: dataframe structured like the proposed data file, has to contain DCT and RE column
* __gene_col__ : name of the gene column
* __sample_col__ : name of the sample column
* __var_col__ : column name of variables to normalize your control against
* __control__: name of variable to use as control

It will pass a dataframe with seven added columns

* __DCTavg__ : average Delta CT value
* __DCTsem__ : s.e.m. of Delta CT
* __DCTsemperc__ : percentage of s.e.m.
* __DDCTavg__ : average Delta Delta CT value
* __DDCTsem__ : standard error to the mean of Delta Delta CT
* __DDCTmin__ : minimum sem value
* __DDCTmax__ : maximum sem value

```{r}
ddct <- calculate_DDCT(df = qpcr, gene_col = "Gene", sample_col = "Sample", var_col = "Sample", control = "WT")
head(ddct)
```

### Plot
```{r}
ggplot(ddct, aes(x = Sample, y = DDCTavg, fill = Sample)) +
  geom_col() +
  geom_errorbar(aes(ymin = DDCTmin, ymax = DDCTmax), width = 0.1, data = ddct) +
  facet_wrap(~Gene, scales = "free_y")
```

# Export
At any given point you can export the data frame to an excel file with this function. Here we will export the data in an excel with three sheets. The first sheet will contain your raw values and normalized delta CT values. The second sheet will contain your statistics. The third sheet will contain all the values normalized to your control group.
```{r}
library(WriteXLS)
WriteXLS(c("qpcr", "stat", "ddct"), "vignette/qpcr.xlsx")
```
