---
title: "Example workflow qPCRanalysis"
author: "Jordi Camps"
date: "26 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r message=FALSE}
library(readxl)
library(tidyr)
library(dplyr)
library(ggplot2)
library(qPCRanalysis)
library(ggpubr)
library(lazyeval)
<<<<<<< HEAD
library(broom)
library(purrr)
library(WriteXLS)
=======
>>>>>>> d7c123854f50fd7e913dceb2b541922fad9c01f1
```

## Load data
We load the data with the **read_excel()** function of the readxl package. The data should consist of at least 4 columns:

* __Sample__ : invidividual names per sample
* __Variable__ : the biological variability in your data set like a disease, different timepoints, perturbation ...
* __Target__ : the gene names of your primers
* __CT__ : the raw CT values

```{r message=FALSE, warning=FALSE}
<<<<<<< HEAD
qpcr <- read_excel("qPCR_data.xlsx", col_names = TRUE, sheet = 3, skip = 35)
head(qpcr)
```

Now we need to clean the data frame. We will select only the columns that we need and change their names because R does not like spaces inside column names.
```{r}
qpcr <- select(qpcr, c("Sample Name", "Target Name", "CT"))
colnames(qpcr) <- c("Sample", "Gene", "CT")
head(qpcr)
```

In this example we want to see the difference between different genotypes. We will split the Sample Name column in two.
```{r}
qpcr <- separate(qpcr, col = Sample, into = c("Sample", "Replicate"), sep = "_")
qpcr <- filter(qpcr, Sample != "Blanc")
head(qpcr)
```

When we check the structure of the data frame we see that the CT column is not numeric, we will change this.
```{r}
qpcr$CT <- as.numeric(qpcr$CT)
str(qpcr)
=======
qpcr <- read_excel("qPCR_data.xlsx", col_names = TRUE)
qpcr$CT <- as.numeric(qpcr$CT)
head(qpcr)
>>>>>>> d7c123854f50fd7e913dceb2b541922fad9c01f1
```

## Calculate Delta CT
To calculate delta CT we use the `calculate_DCT()` function. This function requires four arguments:

* __df__ : dataframe structured like the proposed data file
* __hkg__ : name of housekeeping gene or genes that you want to use to normalize against
* __sample_col__ : name of the sample column
* __gene_col__ : name of the gene column

It will pass a dataframe with three added columns: 

* __CT_hkg__ : average CT value of housekeeping genes
* __DCT__ : Delta CT values
* __RE__ : relative expression to hkg

```{r}
<<<<<<< HEAD
qpcr <- calculate_DCT(df = qpcr, hkg = c("Rab35", "Rpl13a", "PSma3"), sample_col = "Sample", gene_col = "Gene")
head(qpcr)
```

### Plot
```{r}
ggplot(qpcr, aes(x = Sample, y = DCT, col = Sample)) +
  geom_boxplot() +
  facet_wrap(~Gene, scales = "free_y")
```

# Statistics
Perform statistical test between two groups and add information to qpcr data frame.
```{r}
calculate_t.test(qpcr, x = "WT", y = "Sgca")
=======
qpcr <- calculate_DCT(df = qpcr, hkg = c("Rab35", "Rpl13a", "PSma3"), sample_col = "Sample", gene_col = "Target")
>>>>>>> d7c123854f50fd7e913dceb2b541922fad9c01f1
```

## Calculate Delta Delta CT
To calculate Delta Delta CT use the `calculate_DDCT()` function. This function  can only be run after the `calculate_DCT()` function is used and requires five argeuments:

* __df__: dataframe structured like the proposed data file, has to contain DCT and RE column
* __gene_col__ : name of the gene column
* __sample_col__ : name of the sample column
* __var_col__ : column name of variables to normalize your control against
* __control__: name of variable to use as control

It will pass a dataframe with seven added columns

* __DCTavg__ : average Delta CT value
* __DCTsem__ : s.e.m. of Delta CT
* __DCTsemperc__ : percentage of s.e.m.
* __DDCTavg__ : average Delta Delta CT value
* __DDCTsem__ : standard error to the mean of Delta Delta CT
* __DDCTmin__ : minimum sem value
* __DDCTmax__ : maximum sem value

```{r}
<<<<<<< HEAD
qpcr <- calculate_DDCT(df = qpcr, gene_col = "Gene", sample_col = "Sample", var_col = "Sample", control = "WT")
head(qpcr)
```

### Plot
```{r}
ggplot(qpcr, aes(x = Sample, y = DDCTavg, fill = Sample)) +
  geom_col() +
  geom_errorbar(aes(ymin = DDCTmin, ymax = DDCTmax), width = 0.1, data = qpcr) +
  facet_wrap(~Gene, scales = "free_y")
```

# Export
```{r}

```

=======
qpcr <- calculate_DDCT(df = qpcr, gene_col = "Target", sample_col = "Sample", var_col = "Genotype", control = "Healthy")
head(qpcr)
```

>>>>>>> d7c123854f50fd7e913dceb2b541922fad9c01f1
